% latex article template

% cheat sheet(eng): http://www.pvv.ntnu.no/~walle/latex/dokumentasjon/latexsheet.pdf
% cheat sheet2(eng): http://www.pvv.ntnu.no/~walle/latex/dokumentasjon/LaTeX-cheat-sheet.pdf
% reference manual(eng): http://ctan.uib.no/info/latex2e-help-texinfo/latex2e.html

% The document class defines the type of document. Presentation, article, letter, etc. 
\documentclass[12pt, a4paper]{article}

% packages to be used. needed to use images and such things. 
\usepackage[pdfborder=0 0 0]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\PassOptionsToPackage{hyphens}{url}

% hides the section numbering. 
\setcounter{secnumdepth}{-1}

% Graphics/image lications and extensions. 
\DeclareGraphicsExtensions{.pdf, .png, .jpg, .jpeg}
\graphicspath{{./images/}}

% Title or header for the document. 
\title{
	IT3010, Research Methods in IT, Assignment 1. 
}
% Author
\author{
	Magnus L. Kirø \\
}
\date{\today}

\begin{document}
\maketitle
\pagenumbering{arabic}

\begin{abstract}
This is research proposal. We purpose to look into team efficiency and
collaboration in regards to system development and IT. 
%limit to 2 pages. 
\end{abstract}

%idea: research into the performance and structure of system development teams. 

%Write an initial research proposal. The purpose of a research proposal is to
%convince others that you have a worthwhile research project. In this initial
%research proposal, you should include:

\section{Motivation}

\paragraph{problem description}
I propose to look into the issue of team collaboration an effectiveness win
system development. I want to find out what can improve team collaboration and
team self management/ self organization.    

\paragraph{Importance of the problem}
%why is the problem important?.
The problem is important to the development process of information systems, and
other IT-systems.  

\subsection{Relevant articles}
%    a brief review of relevant articles (2-3 articles). 

Articles: 
http://ieeexplore.ieee.org/xpl/articleDetails.jsp?tp=&arnumber=6197202&url=http%3A%2F%2Fieeexplore.ieee.org%2Fxpls%2Fabs_all.jsp%3Farnumber%3D6197202

http://ieeexplore.ieee.org/xpl/articleDetails.jsp?tp=&arnumber=6583162&refinements%3D4291944822%2C4291944246%26ranges%3D2012_2015_p_Publication_Year%26matchBoolean%3Dtrue%26searchField%3DSearch_All_Text%26queryText%3D%28team%2C+system+development%29

http://ieeexplore.ieee.org/xpl/articleDetails.jsp?tp=&arnumber=6613088&refinements%3D4291944822%2C4291944246%26ranges%3D2012_2015_p_Publication_Year%26matchBoolean%3Dtrue%26searchField%3DSearch_All_Text%26queryText%3D%28team%2C+system+development%29

http://ieeexplore.ieee.org/xpl/articleDetails.jsp?tp=&arnumber=6225952&refinements%3D4291944822%2C4291944246%26ranges%3D2012_2015_p_Publication_Year%26matchBoolean%3Dtrue%26searchField%3DSearch_All_Text%26queryText%3D%28team+efficiency%29

% do for all artivcles. 
\paragraph{}
clearly
explaining the main objective and main findings of the article

\paragraph{}
clearly
describing how the study support the findings, 

\paragraph{}
clearly explain how the
article relates to your proposal.


\section{Objectives}
%research objective(s). The research objective(s) should provide a simple, clear and concrete statement of purpose.
The objective of the research is to enlighten the area of system development
and the state of team organization and effectiveness. In detail, looking at how
best to organize and empower development teams in the IT-sector. 

\section{Research Questions}
%research question(s).  The research question(s) should be: 

%(a) answerable,
%(b) consistent with the research objective(s), 
%(c) adequate in term of scope and effort need to address the question(s), 
%(d) related to each other (if there is more than 1 research questions).

\paragraph{1}
What is the state of team organization today?

\paragraph{2}
How can effectiveness be improved in teams today? 

\paragraph{3}
Can empowerment of development teams help efficiency of the development process
and product delivery?

\end{document}
This is never printed
