% latex article template

% cheat sheet(eng): http://www.pvv.ntnu.no/~walle/latex/dokumentasjon/latexsheet.pdf
% cheat sheet2(eng): http://www.pvv.ntnu.no/~walle/latex/dokumentasjon/LaTeX-cheat-sheet.pdf
% reference manual(eng): http://ctan.uib.no/info/latex2e-help-texinfo/latex2e.html

% The document class defines the type of document. Presentation, article, letter, etc. 
\documentclass[12pt, a4paper]{article}

% packages to be used. needed to use images and such things. 
\usepackage[pdfborder=0 0 0]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\PassOptionsToPackage{hyphens}{url}

% hides the section numbering. 
\setcounter{secnumdepth}{-1}

% Graphics/image lications and extensions. 
\DeclareGraphicsExtensions{.pdf, .png, .jpg, .jpeg}
\graphicspath{{./images/}}

% Title or header for the document. 
\title{
	Document Title
}
% Author
\author{
	Magnus L. Kirø \\
	Representing ThisOrThat organization % in what capacity are you presenting this document? as yourself, the sales manager, ceo, etc? 
}
\date{\today}

\begin{document}
\maketitle
\pagenumbering{arabic}

\begin{abstract}
This is the paper's abstract \ldots
\end{abstract}

\section{Task description}
In the assignment, the students are to make a strategic analysis of a real -
world case firm. The strategic analysis is based on the course curriculum.
Thereby, the assignment gives the students experience with applying central
theories, models and concepts from the course. Representatives from the firm
will come twice to provide information about the firm. The first time will be a
lecture - like meeting, with the firm representative(s) providing general
information about the firm, and its current strategic position and challenges.
The second meeting will give the students the opportunity to pose questions to
the firm representative in order to get sufficient information to carry out the
strategic analysis. In addition to these sources of information, it is assumed
that the students make use of publicly accessible information about the firm.
Attendance at the two company lectures is compulsory! The assignment should
contain two main parts:

1.
A strategic analysis of the case company

a.
Mission, values, vision, present goals
and strategies, and strategy processes
b.
Business level strategy, including
i.
Strengths and weaknesses; Inside
-
Out analysis (RBV)
ii.
Opportunities and threats; Outside - In analysis (Porter)

c.
Corporate level strategy, including
i.
Corporate configuration, corporate management, collaboration with sister units,
and history of M&As

d.
Network level strategy, including
i.
The most relevant collaboration partners
, how the com pany cooperates with customers, partners and suppliers

2.
Recommen
ded strategies for the case company

a.
Strategies at business, corporate, and network levels

b.
How the firm’s strategy process
es
should be organized and carried out
If possible, the group who writ es the best assignment will be given the
opportunity to present their results for the case firm. Students may be asked to
sign a confidentiality agreement concerning the information provided by the case
firm. Under no circumstances are students allowed to contact the case company,
or other organizations related to it, directly.

\section{Report format details}
The draft as well as the final report should be of maximum 25 pages, excluding
tables and figures, and the obligatory elements listed below. The number of
words in the main body text should not exceed 10,000. Font size 12 (not narrow
font),2,5 cm indent right/left and top/bottom.

Obligatory elements in the report:

Front page with the group number and the names of all members of the group

Preface

List of contents

List of tables/figures

Abstract

List of references

\section{Norbit Notes}

"explore more"

small entity. 
headquarter in Trh. 

ODM(original design manufactoring), ITS(intelligent traffic systems), SUBSEA,
EMS(electronics manufactoring), 
NORBIT group: 150 employees, MNOK 361 revenue. 

US, Poland, Australia(unmanned atm), associated company in Iceland. 

1995 founded. 
2001 sweeden
2003 denmark
2009 starting to become a corporation. 
2012 US
2014 Poland

Slow growth in the beginning, stronger growth in later years.

vision is important.  

values: 
1. We deliver
2. Safe under pressure
3. Refinement of talents

"Exploring is finding opportunities where others find limits". 
Breaking the values is not acceptable. 

http://vimeo.com/99123303 - presentation of the company. 

bdd, business development director. 

ODM: 
products and systems development. 

...

Knvikna: technical and scientific software. 

:


\end{document}
This is never printed
